Sensitivity Analysis

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Qubic_EOS.tex
21.01

	A cubic equation of state serves as a mathematical model utilized in thermodynamics and fluid dynamics to depict the behavior of real gases and liquids. Its name, "cubic," from its characteristic feature—a third-degree polynomial equation that correlates the pressure, volume, and temperature of a substance. These equations constitute tools for comprehending the phase behavior, properties, and thermodynamic processes of actual substances, across various engineering and scientific applications. The cubic equation of state take into account deviations from ideal gas behavior, which are particularly important at high pressures and low temperatures, where real gases do not follow assumption of ideal gas.

	{\footnotesize
	\begin{equation}
		P = \frac{RT}{v_m-b} - \frac{a\alpha}{v_m^2 - ubv_m + wb^2}
	\end{equation}
	}

	In this equation, P denotes the pressure of the substance, $v_m$ represents the molar volume (volume per mole) of the substance, T stands for the absolute temperature of the substance, u and w are integers that vary from one equation to another, R symbolizes the universal gas constant, and a and b serve as substance-specific parameters known as Van der Waals constants. Additionally, $\omega$ denotes an acentric factor.

	The Van der Waals constants, a and b, constitute empirical values contingent upon the particular substance being modeled. These constants factor in molecular interactions (represented by 'a') and the finite size of gas molecules (indicated by 'b'). Different substances will possess distinct values for 'a' and 'b' based on experimental data.

	Several variations of the cubic equation of state exist, including the Van der Waals equation and the Redlich-Kwong equation, each with its own set of parameters and assumptions. Tables \ref{tab:Popular_Cubic_EoS} show parameters for popular cubic EoS.

@@

	where $A=\frac{a\alpha P}{R^2T^2}$ and $B=\frac{bP}{RT}$.

@@

	For an ideal gas, the compressibility factor is defined as $Z = 1$. In many real-world applications, deviations from ideal gas behavior, needs to be taken into account to have a precise model describing a physical phenomena. The value of $Z$ typically increases with pressure and decreases with temperature. At elevated pressures, molecules collide more frequently, allowing repulsive forces between molecules to influence the molar volume of the real gas ($v_m$) to surpass that of the corresponding ideal gas $\left( \left(v_m\right)_{ideal~gas} = \frac{RT}{P} \right)$, resulting in $Z$ exceeding one. Analogously, at lower pressures, molecules move freely, with attractive forces predominating, leading to $Z < 1$.

	Numerical methods such as Newton-Raphson can be used to solve the polynomial equation to obtain the compressibility ${\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ at given temperature and pressure. Alternatively, the Cardano formula can be used to get the closed form solution of the Equation \ref{EQ:Compressibility_Polynomial}. Further details regarding the Cardano formula can be found in Appendix \ref{CH: Cardano}.

@@

	To calculate the residual contribution, we use the standard equation found in many textbooksr4•6 l for the residual internal energy derived from the Peng-Robinson equation of state

	%Following the work of \citet{Pratt2001}, it is assumed that the specific heat can be obtained as a sum of ideal gas contribution and a residual correction for non-ideal behaviour, as given by Equations \ref{eq: Cv_CO2} and \ref{eq: Cp_CO2}.

	{\footnotesize
		\begin{subequations}\begin{align}
				{\color{black}C_v}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) & = {\color{black}C_v^{\text{I}}}\left({\color{black}T}(t,z),P(t)\right) + {\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) \label{eq: Cv_CO2}; \\ 
				{\color{black}C_p}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) & = \underbrace{{\color{black}C_p^{\text{I}}}\left({\color{black}T}(t,z),P(t)\right)}_{\text{Eq. } \eqref{eq: Cp_I}} + \underbrace{{\color{black}C_p^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}_{\text{Eq. } \eqref{eq: Cp_R}}  \label{eq: Cp_CO2}.
	\end{align}\end{subequations} }

	${\color{black}C_v}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ and ${\color{black}C_p}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ are the specific heat of $CO_2$ at constant volume and pressure, respectively.

	The ideal-gas contribution ${\color{black}C_v^{\text{I}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ and ${\color{black}C_p^{\text{I}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$, can be obtained based on heat-capacity data applicable to gases at low pressure. For $CO_2$, the ideal gas contribution to the specific heat at constant ${\color{black}P}(t)$, as function of ${\color{black}T}(t,z)$ is given by,

	{\footnotesize
		\begin{equation}
			{\color{black}C_p^{\text{I}}}\left({\color{black}T}(t,z),P(t)\right) = {\color{black}C_{P0}} + {\color{black}C_{P1}}T(t,z) + {\color{black}C_{P2}}{\color{black}T}^2(t,z) + {\color{black}C_{P3}}{\color{black}T}^3(t,z) \label{eq: Cp_I}
	\end{equation} }

	The correction term ${\color{black}C_p^{\text{R}}}\left({\color{black}T}(t,z){\color{black}P}(t)\right)$ at constant pressure ${\color{black}P}(t)$, can be defined by introducing  ${\color{black}C_p^{\text{I}}}({\color{black}T}(t,z)) - {\color{black}C_v^{\text{I}}}({\color{black}T}(t,z)) = {\color{black}R}$ and using the definition of ${\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z){\color{black}P}(t)\right)$. The general form of the correction term is given by Equation \ref{eq: Cp_R}

	{\footnotesize	
		\begin{align}
			&{\color{black}C_p^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) = \underbrace{{\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}_{\text{Eq. }\eqref{eq: CvR}} \nonumber \\
			&+ {\color{black}T}(t,z) \underbrace{\left(\cfrac{\partial {\color{black}P}(t)}{\partial {\color{black}T}}\right)_{{\color{black}v_m}(t,z)}}_{\text{Eq. } \eqref{eq: dPdT}} \underbrace{\left(\cfrac{\partial {\color{black}v_m}\left({\color{black}T}(t,z),{\color{black}P}(t)\right)}{\partial {\color{black}T}}\right)_{{\color{black}P}(t)}}_{\text{Eq. } \eqref{eq: dVdT}} - {\color{black}R}. \label{eq: Cp_R} 
	\end{align} }

	%	The braced terms are obtained from the chosen equation of state and the compressibility definition ${\color{black}P}(t) {\color{black}v_m}({\color{black}T}(t,z),{\color{black}P}(t)) = {\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right){\color{black}R}{\color{black}T}(t,z)$. \\

	For the partial derivative of the volume with respect to temperature ${\color{black}T}(t,z)$ at constant pressure ${\color{black}P}(t)$, we have

	{\footnotesize
		\begin{align}
			\left( \cfrac{\partial {\color{black}v_m}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{\partial {\color{black}T}} \right)_{{\color{black}P}(t)} &= \cfrac{{\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) {\color{black}R}}{{\color{black}P}(t)} \nonumber \\
			&+ \cfrac{{\color{black}R}{\color{black}T}(t,z)}{{\color{black}P}(t)} \underbrace{\left( \cfrac{\partial {\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{\partial {\color{black}T}} \right)_{{\color{black}P}(t)}}_{\text{Eq. }\eqref{eq: dZdT}} \label{eq: dVdT}
	\end{align} }

	with partial derivative of the compressibility factor with respect to temperature ${\color{black}T}(t,z)$ at constant pressure ${\color{black}P}(t)$

	{\footnotesize
		\begin{equation}
			\left(\cfrac{\partial{\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{\partial {\color{black}T}}\right)_{{\color{black}P}(t)} = \left(\cfrac{\partial \dfrac{{\color{black}P}(t) {\color{black}v_m}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{{\color{black}R}{\color{black}T}(t,z)}}{\partial {\color{black}T}}\right)_{{\color{black}P}(t)} \label{eq: dZdT}
	\end{equation} }

	Similarly, for the partial derivative of the pressure with respect to temperature at constant volume is :

	{\footnotesize
		\begin{equation}
			\left(\cfrac{\partial {\color{black}P}(t)}{\partial {\color{black}T}}\right)_{{\color{black}v_m}(t,z)} = \left( \cfrac{\partial{\dfrac{{\color{black}Z}{\color{black}R}{\color{black}T}(t,z)}{{\color{black}v_m}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}}}{\partial{\color{black}T}}\right)_{{\color{black}v_m}(t,z)} \label{eq: dPdT}
	\end{equation} }

	The residual specific heat at constant volume is obtained, by definition, by using the residual internal energy

	{\footnotesize
		\begin{equation}
			{\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z),{\color{black}P}(t)\right) = \left(\cfrac{\partial{\color{black}U^{\text{R}}}\left({\color{black}T}(t,z),{\color{black}P}(t)\right)}{\partial {\color{black}T}}\right)_{{\color{black}v_m}(t,z)} \label{eq: CvR}
	\end{equation} }

@@

	In thermodynamics, a departure function is a concept used to calculate the difference between a real fluid's thermodynamic properties and those of an ideal gas, given a specific temperature and pressure. Common departure functions include those for enthalpy, entropy, and internal energy. These functions are used to calculate extensive properties, which are properties computed as a difference between two states.

@@

	The integral in the equation is evaluated over the range of molar volumes from $v_m$ to infinity. The integral includes a term that depends on the derivative of the compressibility factor with respect to temperature, evaluated at the molar volume $v_m$. Finally, the term $1-Z$ is added to account for the deviation of the fluid's properties from those of an ideal gas.

	The Peng–Robinson equation of state relates the three interdependent state properties pressure $P$, temperature $T$, and molar volume $v_m$. From the state properties ($P$, $v_m$, $T$), one may compute the departure function for enthalpy per mole (denoted $h$) as presented by \citet{Gmehling2019} or \citet{Elliott2011}:

>>>>

	A cubic equation of state serves as a mathematical model to describe the behavior of real gases and liquids through a third-degree polynomial equation that correlates the pressure, volume, and temperature of a substance. These equations constitute tools for comprehending the phase behavior, properties, and thermodynamic processes of actual substances, across various engineering and scientific applications. The cubic equation of state take into account deviations from ideal gas behavior, which are particularly important at high pressures and low temperatures, where real gases do not follow assumption of ideal gas.

	{\footnotesize
	\begin{equation}
		P = \frac{RT}{v_m-b} - \frac{\Phi}{v_m^2 - ubv_m + wb^2}
	\end{equation}
	}

	In this equation, P denotes the pressure of the substance, $v_m$ represents the molar volume of the substance, T stands for the absolute temperature of the substance, u and w are integers that vary from one equation to another, R symbolizes the universal gas constant, and a and b serve as substance-specific parameters known as Van der Waals constants. $\omega$ denotes an acentric factor and $\Phi=a\alpha$.

	The Van der Waals constants, a and b, constitute empirical values contingent upon the particular substance being modeled. These constants factor in molecular interactions (represented by 'a') and the finite size of gas molecules (indicated by 'b').

	Several variations of the cubic equation of state exist each with its own set of parameters and assumptions. Tables \ref{tab:Popular_Cubic_EoS} show parameters for popular cubic EoS.

@@

	where $A=\frac{\Phi P}{R^2T^2}$ and $B=\frac{bP}{RT}$.

@@

	For an ideal gas, the compressibility factor is defined as $Z = 1$, but to describe real-life physical phenomena, the deviation of Z needs to be consider. The value of $Z$ typically increases with pressure and decreases with temperature. At elevated pressures, molecules collide more frequently, allowing repulsive forces between molecules to influence the molar volume of the real gas ($v_m$) to surpass that of the corresponding ideal gas $\left( \left(v_m\right)_{ideal~gas} = \frac{RT}{P} \right)$, resulting in $Z$ exceeding one. At lower pressures, molecules move freely, with attractive forces predominating, leading to $Z < 1$.

	Numerical methods such as Newton-Raphson can be used to solve the polynomial equation to obtain the compressibility ${\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ at given temperature and pressure. Alternatively, the closed form solution can by obtained by Cardano formula. Details regarding the Cardano formula can be found in Appendix \ref{CH: Cardano}.

@@

	The residual component of the specific heat can obtained from general relation between $C_v$ and $C_p$ as given by \citet{Poling2001}.

	{\footnotesize
	\begin{equation} \label{EQ:CP_residual}
		C_p^R = C_v^R + T\left(\frac{\partial P}{\partial T}\right)_{v_m} \left(\frac{\partial v_m}{\partial T}\right)_P - R
	\end{equation}}

	The effects of pressure and temperature for liquids are not great, but both $C_p$ and $C_v$ diverge at the critical point of a pure fluid. In the neighborhood of the critical, $\left(\frac{\partial P}{\partial v_m}\right)_T$ approaches zero, so $C_p$ increases much faster than $C_v$. At both high and low densities, the differences are small, but for $T_r$ near unity, they increase rapidly as the critical density is approached. At fixed density in this region, $C_p$ actually decreases as T increases.

	The term $\left(\frac{\partial P}{\partial T}\right)_{v_m}$ can be obtain by direct differentiation of P-R EoS

	{\footnotesize
	\begin{equation}
		\left(\frac{\partial P}{\partial T}\right)_{v_m} = \frac{R}{v_m - b} - \frac{\frac{d \Phi}{dT}}{\left[v_m(v_m+b)+b(v_m-b)\right]^2}
	\end{equation}}

	where $\left(\frac{\partial \Phi}{\partial T}\right)_{v_m} = \frac{-f(\omega)a}{\sqrt{TT_c}(1 + f(\omega)( 1 - \sqrt{T/T_c}))}$

	The term $\left(\frac{\partial v_m}{\partial T}\right)_P = \frac{R}{P} \left(T \left(\frac{\partial Z}{\partial T}\right)_P + Z \right)$

	The term $\left(\frac{\partial Z}{\partial T}\right)_P$ can be defined as below

	{\footnotesize
	\begin{equation}
		\left(\frac{\partial Z}{\partial T}\right)_P = \frac{ \left(\frac{\partial A}{\partial T}\right)_P (B-Z) + \left( \frac{\partial B}{\partial T} \right)_P \left( 6BZ+2Z-3B^2-2B+A-Z^2 \right)}{3Z^2 + 2(B-1)Z + (A-2B-3B^2)}
	\end{equation}}

	{\footnotesize
	\begin{align}
		\left(\frac{\partial A}{\partial T}\right)_P &= (P/(RT)^2)(\frac{d \Phi}{dT} - 2a/T) \\
		\left(\frac{\partial B}{\partial T}\right)_P &= -bP/(RT^2)
	\end{align}}

	The first component of Equation \ref{EQ:CP_residual} from definition of $C_v = \left(\frac{\partial U^R}{\partial T}\right)_{v_m}$, where $U^R$ represents internal energy. In case of Pen-Robinson $U^R$ equals to

	{\footnotesize
	\begin{equation}
		U^R = \frac{T \frac{d^2 \Phi}{dT^2}}{b \sqrt{8}} \ln \left( \frac{Z+B\left(1+\sqrt{2}\right)}{Z+B\left(1-\sqrt{2}\right)} \right)
	\end{equation}}

	As given by \citet{Pratt2001}, for Peng-Robinson EoS the term $\frac{d^2 \Phi}{dT^2}$ is defined as follow

	{\footnotesize
	\begin{equation}
		\frac{d^2\Phi}{dT^2} = \frac{af(\omega)\sqrt{\frac{T_c}{T}} \left( 1+ f\left(\omega\right) \right)}{2TT_c}
	\end{equation} }

@@

	In thermodynamics, a departure function is a concept used to calculate the difference between a real fluid's thermodynamic properties (enthalpy, entropy, or internal energy) and those of an ideal gas, given a specific temperature and pressure. These functions are used to calculate extensive properties, which are properties computed as a difference between two states.

@@

	The integral in the equation is evaluated over the range of molar volumes from $v_m$ to infinity. The integral includes a term that depends on the derivative of the compressibility factor with respect to temperature, evaluated at the molar volume $v_m$. Finally, the term $1-Z$ is added to account for the deviation of the fluid's properties from an ideal state.

	The Peng–Robinson EoS relates the three interdependent state properties pressure $P$, temperature $T$, and molar volume $v_m$. From the state properties ($P$, $v_m$, $T$), one may compute the departure function for enthalpy per mole (denoted $h$) as presented by \citet{Gmehling2019} or \citet{Elliott2011}:

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Results.tex
21.01

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
			\csvautobooktabular[separator=comma, table head = \toprule & $40~^\circ C$ & $50~^\circ C$ & $40~^\circ C$ & $50~^\circ C$ \\ & $200~bar$ & $200~bar$ & $300~bar$ & $300~bar$ \\\midrule]{Figures/Results_estimation/estimation.csv} }
		\caption{Parameter estimation results}
		\label{tab: Estimation_results}
	\end{table}

@@

		\begin{figure}[!h]
		\centering
		\begin{subfigure}[b]{0.49\columnwidth}
			\centering
			\includegraphics[trim = 10.5cm 14.0cm 10.0cm 1.0cm,clip,width=\columnwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
			\caption{Dependency of ${\color{black}D_i}$}
			\label{fig: results_DI}
		\end{subfigure}
		\begin{subfigure}[b]{0.49\columnwidth}
			\centering
			\includegraphics[trim = 3.0cm 7.9cm 17.7cm 7.7cm,clip,width=\columnwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
			\caption{Dependency of ${\color{black}\Upsilon}$}
			\label{fig: results_upsilon}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.49\columnwidth}
			\includegraphics[trim = 1.0cm 0cm 2.0cm 0cm, clip, width=\columnwidth]{/Results_estimation/D_i_Uppsilon_results.pdf}
			\caption{Dependency of ${\color{black}D^R_i}$}
			\label{fig:Di_upsilon_res}
		\end{subfigure}
		\begin{subfigure}[b]{0.49\columnwidth}
			\centering
			\includegraphics[trim = 18.0cm 14.0cm 2.5cm 1.0cm,clip,width=\columnwidth]{/Results_estimation/Trend_Lines_RE_order_1.pdf}
			\caption{Dependency of ${\color{black}D^M_e}$}
			\label{fig: results_De}
		\end{subfigure}
		\caption{Results of parameter estimation}
		\label{fig: estimation_results_DI_GAMMA}
	\end{figure}

>>>>

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
			\csvautobooktabular[separator=comma, table head = \toprule & $40~^\circ C$ & $50~^\circ C$ & $40~^\circ C$ & $50~^\circ C$ \\ & $200~bar$ & $200~bar$ & $300~bar$ & $300~bar$ \\\midrule]{Figures/Results_estimation/estimation_2.csv} }
		\caption{Parameter estimation results}
		\label{tab: Estimation_results}
	\end{table}

@@

		\begin{figure}[!h]
		\centering
		\begin{subfigure}[b]{0.49\columnwidth}
			\centering
			\includegraphics[trim = 11.2cm 13.7cm 11cm 1.0cm,clip,width=\columnwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
			\caption{Dependency of ${\color{black}D_i}$}
			\label{fig: results_DI}
		\end{subfigure}
		\begin{subfigure}[b]{0.49\columnwidth}
			\centering
			\includegraphics[trim = 3.5cm 7.7cm 18.7cm 7.3cm,clip,width=\columnwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
			\caption{Dependency of ${\color{black}\Upsilon}$}
			\label{fig: results_upsilon}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.49\columnwidth}
			\includegraphics[trim = 1.0cm 0cm 3cm 0cm, clip, width=\columnwidth]{/Results_estimation/D_i_Uppsilon_results.pdf}
			\caption{Dependency of ${\color{black}D^R_i}$}
			\label{fig:Di_upsilon_res}
		\end{subfigure}
		\begin{subfigure}[b]{0.49\columnwidth}
			\centering
			\includegraphics[trim = 19.0cm 13.8cm 3.5cm 1.0cm,clip,width=\columnwidth]{/Results_estimation/Trend_Lines_D_e.pdf}
			\caption{Dependency of ${\color{black}D^M_e}$}
			\label{fig: results_De}
		\end{subfigure}
		\caption{Results of parameter estimation}
		\label{fig: estimation_results_DI_GAMMA}
	\end{figure}

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Sensitivity_Analysis.tex
21.01

	As presented in the work of \citet{Maly1996}, the sensitivity analysis equations(${\dot{Z}}$) are developed by taking the total derivative of the state vector $x$ with respect to parameters $p$, where $p$ is a subset of $\theta$.

@@

	The sensitivity matrix ${S}({x}(t);{p})$ represents the matrix of equations of size $N_x \times N_p$, where each equation ${S}(n_x,n_p)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $p_{n_p}$.

@@

	The jacobian ${J_p}({x}(t);{p})$ represents the matrix of equations of size $N_x \times N_p$, where each equation ${J_p}(n_x,n_p)$ is the partial derivative of the process model equation $F_{n_x}$ with respect to the parameter $p_{n_p}$.

>>>>

	As presented in the work of \citet{Maly1996}, the sensitivity analysis equations(${\dot{Z}}$) are developed by taking the total derivative of the state vector $x$ with respect to parameters $p$, where $p$ is a subset of the parameter space $\theta$.

@@

	The sensitivity matrix ${S}({x}(t);{p})$ represents the matrix of equations of size $N_x \times N_p$, where each subequation ${S}(n_x,n_p)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $p_{n_p}$.

@@

	The jacobian ${J_p}({x}(t);{p})$ represents the matrix of equations of size $N_x \times N_p$, where each subequation ${J_p}(n_x,n_p)$ is the partial derivative of the process model equation $F_{n_x}$ with respect to the parameter $p_{n_p}$.

### 2024-01-21 SFE_article_Sensitivity_Analysis/Article_Sensitivity_Analsysis.tex
21.01

The sensitivity analysis is a tool to understand the parametric dependence of dynamic behavior and its suitability in inferring mechanisms of the dynamic behavior of the analyzed system. The mathematical formulation involves derivative-based local sensitivity analysis of the solution of an inverse problem with respect to perturbations of parameters. The local method is based only on considering all parameters as constants (those independent of time or location), and then the sensitivity equations are estimated by solving a variational system simultaneously with the original system. The sensitivity equations can be obtained in various ways, but the automatic differentiation technique was implemented to obtain the sensitivity equations in this work. By identifying which parameters are influential at each period of the simulation, local sensitivity analysis can yield valuable information for guiding the process modelling, the design of experiments or the model reduction. The local sensitivity analysis techniques consider the sensitivities at only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear system.

In this study, the local sensitivity analysis method was introduced to explore the dynamics of the supercritical extraction system, which consists of a set of partial differential equations. The sensitivity equations evaluate the influence of control variables such as flow rate, pressure, and inlet temperature on the state-space consisting of the concentration of the solute in solid and liquid phases, the fluid's enthalpy density, system pressure, and extraction yield. Every change in the control variables leads to a state-space change, eventually affecting the process performance. This information can be utilized to identify which control variables influence the extraction yield the most. The controls with high sensitivities on the extraction can be used to investigate to find optimal operating conditions from an economic point of view.

>>>>

The sensitivity analysis is a tool to understand the parametric dependence in the dynamic behaviour of the analysed system. The presented formulation involves derivative-based local sensitivity analysis of the model solution with respect to selected parameters and controls. The sensitivity equations can be obtained in various ways, but the automatic differentiation technique was implemented to obtain the sensitivity equations in this work. By identifying which parameters are influential at each period of the simulation, local sensitivity analysis can yield valuable information for guiding the process modelling, the design of experiments or the model reduction. The local sensitivity analysis techniques consider the sensitivities at only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear model.

In this study, the local sensitivity analysis method was introduced to explore the dynamics of the supercritical extraction system, which consists of a set of partial differential equations. The sensitivity equations evaluate the influence of control variables such as flow rate, pressure, and inlet temperature on the state-space consisting of the concentration of the solute in solid and liquid phases, the fluid's enthalpy density, system pressure, and extraction yield. Every change in the control variables leads to a state-space change, eventually affecting the process performance.

Similarities between Figures \ref{fig:Sensitivty_F_y}, \ref{fig:Sensitivty_P_y} and \ref{fig:Sensitivty_T_y} can observed, although the sources of disturbances are different. All the plots stay initially plat, and this is because the extraction kinetic is mainly controlled by the concentration gradient. The largest sensitivity deviations are observed in the second extraction phase, which is mainly controlled by the internal mass transfer. These deviations are strongly related to the fluid velocity and the correlations used to connect the extraction kinetic parameters. As correlations used in this work are linear functions of the fluid density( {\color{red} article 1}), Figures \ref{fig:Sensitivty_P_y} and \ref{fig:Sensitivty_T_y} are characterised by the opposite changes in the fluid density, which leads to the sensitivity plots with similar shape but opposite sign. In the third extraction phase, the concentration gradient and the internal diffusion coefficient decrease with the solute concentration in the solid phase( it is more difficult to extract the solute from the particle core than from the outer shell). It is important to note that the local-sensitivity methods obtain the discussed results, which can be obtained at different operating conditions.

This information can be utilized to identify which control variables influence the extraction yield the most. The controls with high sensitivities on the extraction can be used to investigate to find optimal operating conditions from an economic point of view.
